---
output: bookdown::pdf_document2
---

# Talker analysis: Characterizing voiceless stop VOT distributions {#explore-spk}

In this section, we analyze the VOT distributions of the talkers.
Table \@ref(tab:spk-vot-tab) provides descriptive statistics of the VOT distributions for both voiced and voiceless stops by task, phoneme, and talker.

```{r spk-vot-tab}
avg_voice %>%
  kable(col.names = c("Phase", "Onset", "\\textit{N}", rep(c("\\textit{M}", "\\textit{SD}", "Range"), 4)),
        caption = "Mean, standard deviation, and range of VOTs by onset by talker.", 
        escape = FALSE) %>%
  kable_styling(bootstrap_options = "bordered", latex_options = "scale_down") %>%
  add_header_above(header = c(" " = 3, "Talker 1" = 3, "Talker 2" = 3, "Talker 3" = 3, "Talker 4" = 3)) %>%
  collapse_rows(columns = 1)
```

## Talker-specific analysis of voiceless stop VOT

We first investigate the extent to which each talker's voiceless stop VOTs diverge from L1 US English "norms," which represent our listeners' prior beliefs about these distributions.
We used the VOT dataset compiled by @chodroff2019 to determine the mean VOT for each voiceless stop onset among L1 US English talkers (see Table \@ref(tab:spk-vot-mu)).
We then conducted one-sample t-tests to compare each of the talkers' mean VOTs to the norm.
These analyses included both the exposure (multisyllabic) and test (monosyllabic) stimuli (48 items per onset per talker).
Table \@ref(tab:spk-vot-mu) provides the difference in means and asterisks indicating the significance level from each analysis (corrected for multiple comparisons with the Hommel method).

```{r spk-vot-mu}
tab_voice_mu %>%
  kable(caption = "Mean difference between L1 US English and talker mean VOTs by onset. Asterisks indicate significance levels from two-sample t-tests: *** < .001, ** < .01, * < .05. English means from Chodroff and Wilson (2019).",
        escape = FALSE) %>%
  kable_styling(bootstrap_options = "bordered", latex_options = "scale_down") %>%
  collapse_rows(columns = 1)
```

## Between-talker analysis of voiceless stop VOT

We next investigate the extent to which each talker's mean VOTs differed from the other talkers'.
This analysis provides a point of comparison to Xie and Myers' (2017) operationalization of exposure-test similarity.
Here, we conducted paired two-sample t-tests on each combination of talkers by onset (48 items per onset per talker).
Table \@ref(tab:spk-vot-mat) shows the results of these comparisons, with the value in each cell being the column talker's mean minus the row talker's mean.

```{r spk-vot-mat}
matrix_voice %>%
  kable(caption = "Mean difference in VOT between talkers by onset. Asterisks indicate significance levels from paired two-sample t-tests corrected with Hommel method: *** < .001, ** < .01, * < .05.",
        escape = FALSE) %>%
  kable_styling(bootstrap_options = "bordered", latex_options = "scale_down") %>%
  collapse_rows(columns = 1)
```

## Discussion

The two analyses conducted in this section show that our talkers differed from both L1 US English norms and one another in their VOT distributions.
Talker 4 produced the most Spanish-like VOT distributions across places of articulation, with significantly shorter means in all comparisons.
Talker 3 also differed from the L1 US English norms and the three other talkers, but in the opposite direction: her mean VOTs were significantly longer in all comparisons.
Talkers 1 and 2 were the most similar to one another, with statistically equivalent mean VOTs for /p/ and /k/.
Talker 1 was the most similar to the L1 US English norms, though both she and Talker 2 had significantly longer /k/ VOTs.
Overall, each talker exhibited a unique pattern of VOT-stop distributions.
We discuss the implications of these results for our Similarity and Variability manipulations in the next section.
